Presentation Title

### 2017-07-30 thesis/chapters/ch3-model.tex
Edit ch3, 4 and 5

We present a design which removes the transaction rate restrictions seen in traditional blockchain systems.
This is done by combining the recent HoneyBadgerBFT~\cite{miller2016honey}
(a purely asynchronous consensus algorithm for blockchain systems)
with a modification of our prior work on TrustChain~\cite{trustchain}.
The design is inspired by (the distributed part of) the Internet architecture as well as how transactions are performed in the real world.

@@

\section{System overview}
\label{sec:system-overview}
The system consist of one data structure---Extended TrustChain,
and three protocols---consensus protocol, transaction protocol and validation protocol.
We first describe each component individually and then explain how they fit together in \Cref{sec:combined-protocol}.

@@

\subsubsection*{Standard TrustChain}
In TrustChain, every node has a ``personal'' chain. 
Initially, the chain only contains a genesis block generated by the nodes themselves.
When a node $A$ wishes to add a new transaction (TX) with $B$, a new TX block is generated and then appended to $A$'s chain.
The TX block must have a valid hash pointer pointing to the previous block
and a reference\footnote{This is different from the original TrustChain definition found in~\cite{trustchain}.
In there, a TX block has two outgoing edges which are hash pointers to the two parties involved in the transaction.
This work uses one outgoing edge and a reference.} to its \emph{pair} on $B$'s chain.
As a result, a single transaction generates two TX blocks, one on each party's chain.
An example of is shown in \Cref{fig:trustchain-bad}.

@@

If every node follows the rules of TrustChain and we only consider hash pointers,
then every chain effectively forms a singly linked list.
However, if a node violates the rules, then a \emph{fork} may happen.
That is, there may be more than one TX block with a hash pointer pointing back to the same block.
In \Cref{fig:trustchain-bad}, node $b$ (in the middle chain) created two TX blocks that both point to $t_{b, 5}$.
If this is a ledger system it can be seen as a double spend, where the currency accumulated up until $t_{b, 5}$ are spent twice.

@@

\subsection{Consensus protocol}\label{sec:overview-cons}

@@

The high communication complexity of Byzantine consensus algorithms prohibits us from running it on a large network.
Thus, for every round, we randomly select some node---called facilitators---to collect CP blocks from every other node and use them as the proposal.
The facilitators are elected using a \emph{luck value}, which is computed using $\textsf{H}(\C_r || pk_u)$,
where $\C_r$ is the consensus result 
(which can be seen as the set union of all the CP blocks collected by the facilitators)
in round $r$ and $pk_u$ is the public key of node $u$.
Intuitively, the election is guaranteed to be random 
because the output of a cryptographically secure hash function is unpredictable and $\C_r$ cannot be determined in advance.

@@

The consensus and transaction protocol by themselves do not provide a mechanism to detect forks or other forms of tamperaing.
Thus we need a validation protocol to counteract malicious behaviour.
When a node wish to validate one of its TX, it asks the counterparty for the \emph{fragment} of the TX.
A fragment of a TX is a section of the chain beginning and ending with CP blocks that contains the TX and are in consensus.
Upon the counterparty's response, the node checks that the CP blocks are indeed in consensus,
the hash pointers are valid and his TX is actually in the fragment.
The TX is valid if these conditions are satisfied.
Intuitively, this works because it is hard (because a cryptographically secure hash function is second-preimage resistant)
to create a different chain that begins and ends with the same two CP blocks but with a different middle section.

@@

This section and the ones following it give a technical treatment of what the content in System Overview.
For notational clarity, we use the following convention (adapted from~\cite{miller2016honey}) throughout this work.
\begin{itemize}
\item Lower case (e.g. $x$) denotes a scalar object or a tuple.
\item Upper case (e.g. $X$) denotes a set or a constant.
\item Sans serif (e.g. $\textsf{fn}(\cdot)$) denotes a function.
\item Monospace (e.g. $\texttt{ack}$) denotes message type.
\end{itemize}
Further, we use $a || b$ to denote concatination of the binary representations of $a$ and $b$.

We assume purely asynchronous channels with eventual delivery.
Thus in no circumstance do we make timing assumptions.
The adversary has fully control of the delivery schedule and the message ordering of all messages.
But they are not allowed to drop messages except for their own\footnote{
    Reliability can be achieved in unreliable networks by resending messages or using some error correction code.
}.

We assume there exist a Public Key Infrastructure (PKI), and nodes are identified by their unique and permanent public key.
Finally, we use the random oracle model, i.e. calls to the random oracle are denoted by $\textsf{H}: \{0, 1\}^* \rightarrow \{0, 1\}^\lambda$,
where $\{0, 1\}^*$ denotes the space of finite binary strings and $\lambda$ is the security parameter~\cite{bellare1993random}.

In our model we consider $N$ nodes, which is the population size.
$n$ of them are facilitators, $t$ out of $n$ are malicious and the inequality
$n \ge 3t + 1$ must hold.
This is from the work of Pease, Shostak and Lamport, where they show a network of $n$ nodes cannot reach Byzantine agreement with $t \ge n/3$.
Further, the inequality $N \ge n + t$ must also hold.
This is due to our system design, which becomes clear in~\Cref{sec:consensus-phase}.

Our threat model is as follows. 
We use a restricted version of the adaptive corruption model.
The first restriction is that corrupted node can only change across rounds.
That is, if a round has started, the corrupted nodes cannot changed until the next round.
The second restriction is that the adversary, presumably controlling all the corrupted nodes, is forgetful.
Namely the adversary may learn the internal state such as the private key of a corrupted node,
but if the node recovers, then the adversary must forget the private key.
This is realistic because otherwise the adversary can eventually learn all the private keys and sabotage the system.
Finally, we assume computational security.
That is, the adversary can run polynomial-time algorithms but not exponential-time algorithms efficiently.

\section{Extended TrustChain}
The primary data structure used in our system is the Extended TrustChain.
Each node $u$ has a public and private key pair---$pk_u$ and $sk_u$, and a chain $B_u$.
The chain consist of blocks $B_u = \{ b_{u, i} : k \in \{ 0, \dots, h - 1 \} \}$,
where $b_{u, i}$ is the $i$th block of $u$,
and $h = |B_u|$.
We often use $b_{u, h}$ to denote the latest block.
There are two types of blocks, TX blocks and CP blocks.
If $T_u$ is the set of TX blocks of $u$ and $C_u$ is the set of CP blocks of $u$,
then it must be the case that $T_u \cup C_u = B_u$ and $T_u \cap C_u = \varnothing$.
The notation $b_{u, i}$ is generic over the block type.
We assume there exist a function $\textsf{typeof}: B_u \rightarrow \{ \tau, \gamma \}$ that returns the type of the block,
where $\tau$ represents the TX type and $\gamma$ represents the CP type.

\subsection{Transaction block}
The TX block is a six-tuple, i.e $t_{u, i} = \langle \textsf{H}(b_{u, i - 1}), seq_u, txid, pk_v, m, sig_u \rangle$.
We describe each item in turn.
\begin{enumerate}
\item $\textsf{H}(b_{u, i - 1})$ is the hash pointer to the previous block.
\item $seq_u$ is the sequence number which should equal $i$.
\item $txid$ is the transaction identifier, it should be generated using a cryptographically secure pseudo-random number generator by the initiator of the transaction.
\item $pk_v$ is the public key of the counterparty $v$.
\item $m$ is the transaction message.
\item $sig_u$ is the signature created using $sk_u$ on the concatination of the binary representation of the five items above.
\end{enumerate}
The fact that we have no constraint on the content of $m$ is in alignment with our design goal---application neutrality.

TX blocks come in pairs.
In particular, for every block 
$$t_{u, i} = \langle \textsf{H}(b_{u, i - 1}), seq_u, txid, pk_v, m, sig_u \rangle$$
there exist one and only one \emph{pair} 
$$t_{v, j} = \langle \textsf{H}(b_{v, j - 1}), seq_v, txid, pk_u, m, sig_v \rangle.$$
Note that the $txid$ and $m$ are the same, and the public keys refer to each other.
Thus, given a TX block, these properties allow us to identify its pair.

@@

\subsection{Checkpoint block}

The CP block is a five-tuple, 
i.e. $c_{u, i} = \langle \textsf{H}(b_{u, i-1}), seq_u, \textsf{H}(\C_r), r, sig_u \rangle$,
where $\C_r$ is the consensus result (which we describe in \Cref{sec:consensus-result}) in round $r$, the other items are the same as the TX block definition.
Note that unlike in our prior work~\cite{implicitconsensus}, CP blocks and TX blocks do not have independent sequence numbers.

The genesis block in the chain must be a CP block in the form of
$c_{u, 0} = \langle \textsf{H}(\bot), 0,  \textsf{H}(\bot), 0, sig_u \rangle$
where $\textsf{H}(\bot)$ can be interpreted as applying the hash function on an empty string.
The genesis block is unique because every node has a unique public and private key pair.

\subsection{Consensus result}
\label{sec:consensus-result}
Our consensus protocol runs in rounds as discussed in \Cref{sec:system-overview}.
Every round is identified by a round number $r$, which is incremented on every new round.
The consensus result is a tuple, i.e. $\C_r = \langle r, C \rangle$,
where $C$ is a set of CP blocks agreed by the facilitators of round $r$.

\subsection{Chain properties}
Here we define a few important properties which results from the interleaving nature of CP and TX blocks.

If there exist a tuple $\langle c_{u,a}, c_{u, b} \rangle$ for a TX block $t_{u, i}$,
where 
$$a = \argmin_{k, k < i, \texttt{typeof}(b_{u,k}) = \gamma}(i - k)$$
$$b = \argmin_{k, k > i, \texttt{typeof}(b_{u,k}) = \gamma}(k - i),$$
then $\langle c_{u,a}, c_{u, b} \rangle$ is the \emph{enclosure} of $t_{u, i}$.
Note that $c_{u, a}$ is the more recent CP block.
Also, some TX blocks may not have any subsequent CP blocks, then its enclosure is $\bot$.

If the enclosure of some TX block $t_{u, i}$ is $\langle c_{u,a}, c_{u, b} \rangle$,
then its \emph{fragment} $F_{u, i}$ is defined as $\{ b_{u, i} : a \le i \le b \}$.
For convenience, the function $\textsf{fragment}(\cdot)$ represents the fragment of some TX block if it exists, otherwise $\bot$.

\emph{Agreed enclosure} is the same as enclosure with an extra constraint where the CP blocks must be in some consensus result $\C_r$ and also must be the smallest enclosure.
That is, suppose a chain is in the form
    $\{c_{i}, c_{i+1}, t_{i+2}, c_{i+3}\}$\footnote{Usually the notation is of the form $c_{u, i}$, but the node identity is not important here so we simplify it to $c_{i}$}
    and $c_{i}, c_{i+1}, c_{i+3}$ are in $\C_r, \C_{r+1}, \C_{r+3}$ respectively,
    then the agreed enclosure of $t_{i+2}$ is $\langle c_{i+1}, c_{i+3}\rangle$ and cannot be $\langle c_{i}, c_{i+3}\rangle$.
Similarly, \emph{agreed fragment} has the same definition as fragment but using agreed enclosure.
We define its function to be $\textsf{agreed\_fragment}(\cdot)$ which we use later in the validation protocol (\Cref{sec:vd-protocol}).

@@

Our consensus protocol runs on top of Extended TrustChain.
It is directly related to the creation of CP blocks.
The objectives of the protocol are to
    allow honest nodes always make progress (in the form of creating new CP blocks),
    compute correct consensus result in every round
    and have unbiased election of facilitators.
Concretely, we define the necessary properties as follows.
\begin{definition}
\label{def:consensus}
\textbf{\emph{Properties of the consensus protocol}}

$\forall r \in \mathbb{N}$, the following properties must hold.
\begin{enumerate}
    \item \emph{Agreement}:
        If one correct node outputs a set of facilitators $\F_r$,
        then every node outputs $\F_r$
    \item \emph{Validity}:
        If any correct node outputs $\F_r$, then 
            \begin{enumerate}
                \item $|\C_r| \ge N - t$ must hold for the $\C_r$ which was used to create $\F_r$,
                \item $\F_r$ must contain at least $n - t$ honest nodes and
                \item $|\F_r| = n$.
            \end{enumerate}
    \item \emph{Fairness}:
        Every node with a CP block in $\C_r$ should have an equal probability of becoming a member of $\F_r$.
    \item \emph{Termination}:
        Every correct node eventually outputs some $\F_r$.
\end{enumerate}
\end{definition}
These properties may look like Byzantine consensus properties (which we describe next in \Cref{sec:acs-background}) but they have some subtle differences.
Firstly, they are properties for every node in the network and not just the facilitators.
Secondly, they must be satisfied for all rounds because the whole system falls apart if one of the property cannot be satisfied in one of the rounds.

Before describing the protocol in detail,
we take a brief detour to give background on the asynchronous subset consensus.
This is the primary building block of our protocol.
% Starting with the bootstrap phase and then moving on to the actual consensus phase.

@@

The best way to explain asynchronous subset consensus (ACS) is to contrast it with the typical Byzantine consensus.
We adapt the description from~\cite[Chapter 17]{podc}.
%The Byzantine consensus problem is a classical problem is distributed systems literature.
%It is first described by Leslie Lamport as the Byzantine general's problem in 1982~\cite{lamport1982byzantine}.
%Byzantine consensus is described as follows (adapted from ).
\begin{definition}
\textbf{\emph{Byzantine Consensus}}

@@

ACS shares many similarities with Byzantine consensus.
But it is an especially useful primitive for blockchain systems.
It allows any party to propose a value and the result is the set union of all the proposed values by the majority.
This is the primary difference with Byzantine consensus.
Concretely, ACS needs to satisfy the following properties (adapted from~\cite{miller2016honey}).
\begin{definition}
\label{def:acs}
\textbf{\emph{Asynchronous Subset Consensus}}

@@

ACS has the nice property of censorship resiliance when compared to other consensus algorithms.
For instance, Hyperledger and Tendermint uses Practical Byzantine Fault Tolerance (PBFT)~\cite{castro1999practical} as their consensus algorithm.
In PBFT, a leader is elected, if the leader is malicious but follows the protocol, then it can selectivly filter transactions.
In contract, every party in ACS are involved in the proposal phase,
and it guarantees that if $n - 2t$ parties propose the same transaction, then it must be in the agreed output.
Thus, if some value is submitted to at least $n - 2t$ nodes, it is guaranteed to be in the consensus result.
For a detailed description of ACS we refer to the HoneyBadgerBFT work~\cite{miller2016honey}.

@@

\subsection{Bootstrap phase}
\label{sec:bootstrap}
Now we have all the necessary information to describe our consensus protocol.
We begin with the bootstrap phase and then move onto the actual consensus phase.

Recall that facilitators are computed from the consensus result,
but the consensus result is agreed by the facilitators.
Thus we have a dependency cycle.
The goal of the bootstrap phase is to give us a starting point in the cycle.

To bootstrap, imagine that there is some bootstrap oracle, that initiates the code on every node.
The code satisfied all the properties in \Cref{def:consensus}.
Namely every node has the same set of valid facilitators $\F_1$ that are randomly choosen.
This concludes the bootstrap phase.

@@

\subsection{Consensus phase}
\label{sec:consensus-phase}
The consensus phase begins when $\F_r$ is available to all the nodes.
Note that $\F_r$ indicates the facilitators that were elected using result of round $r$ and are responsible for driving the ACS protocol in round $r + 1$.
The goal is to reach agreement on a set of new facilitators $\F_{r + 1}$ that satisfies the four properties in \Cref{def:consensus}.

There are two scenarios in the consensus phase.
First, if node $u$ is not the facilitator, it sends $\langle \texttt{cp\_msg}, c_{u, h} \rangle$ to all the facilitators.
Second, if the node is a facilitator, it waits until it has received $N - t$ messages of type \texttt{cp\_msg}.
% Second if the node is a facilitator, it waits for some duration $D$ and collect messages of type \texttt{cp\_msg}, where $D \gg \Delta$.
Invalid messages are removed, namely blocks with invalid signatures and duplicate blocks signed by the same key.
With the sufficient number of \texttt{cp\_msg} messages,
it begins the ACS algorithm and some $\C'_{r + 1}$ should be agreed upon by the end of it.
Duplicates and blocks with invalid signatures are again removed from $\C'_{r+1}$ and we call the final result $\C_{r+1}$
We have the remove invalid blocks a second time (after ACS) because the adversary may send different CP blocks to different facilitators,
which results in invalid blocks in the ACS output.

At the core of the consensus phase is the ACS protocol.
While any ACS protocol that satisfies the standard definition will work,
we use a simplification of HoneyBadgerBFT~\cite{miller2016honey} as our ACS protocol
because it is the only (to the best of our knowledge) consensus algorithms designed for blockchain systems.
We do not use the full HoneyBadgerBFT due to the following.
First, the transactions in HoneyBadgerBFT are first queued in a buffer and the main consensus algorithm starts only when the buffer reaches an optimal size.
We do not have an infinite stream of CP blocks, thus buffering is unsuitable.
Second, HoneyBadgerBFT uses threshold encryption to hide the content of the transactions.
But we do not reach consensus on transactions, only CP blocks, so hiding CP blocks is meaningless for us as it contains no transactional information.

@@

Upon receiving $\C_{r + 1}$ and at least $n - f$ valid signatures by some node $u$, $u$ performs two asks.
First, it creates a new CP block using $\textsf{new\_cp}(\C_{r + 1}, r + 1)$ (\Cref{alg:new-cp}).
Second, it computes the new facilitators using $\textsf{get\_facilitator}(C, n)$ (\Cref{alg:facilitator})
and updates its facilitator set to the result.
This concludes the consensus phase and brings us back to the beginning of the consensus phase.

\begin{algorithm}
\caption{Function $\textsf{get\_facilitator}(C, n)$ takes a set of CP blocks $C$ and an integer $n$,
sort evey element in $C$ by its luck value (the $\lambda$-expression), and outputs the smallest $n$ elements.}
\label{alg:facilitator}
\begin{algorithmic}
\State $\textsf{take} (n, \textsf{sort\_by} (\lambda x.\textsf{H}(x || pk \text{ of } x), C))$
\end{algorithmic}
\end{algorithm}

>>>>

We design a novel system in order to achieve our ambitious goals outline in~\Cref{ch:problem}.
The system consist of three protocols---the consensus protocol, the transaction protocol and the validation protocol.
It is based on our prior work on TrustChain~\cite{trustchain}, where every node independently interact with each other via their own blockchain.

@@

\section{System overview}
\label{sec:system-overview}
The architecture of our system is visualised in~\Cref{fig:architecture}.
\begin{figure}[h]
\centering
\includegraphics[width=0.7\textwidth]{architecture}
\caption{Overview of the system architecture}
\label{fig:architecture}
\end{figure}
It consist of one data structure---Extended TrustChain, and three protocols.
The protocols have distinct roles which are evident from their names.
They run concurrently and do not interact with each other.
The only synchronisation happens on the Extended TrustChain layer, where all the protocols read and write from it.
We first describe each component individually, starting with Extended TrustChain, and then explain how they fit together.

@@

\subsubsection*{Standard TrustChain}
In TrustChain, every node has a ``personal'' chain. 
Initially, the chain only contains a genesis block generated by the nodes themselves.
When a node $A$ wishes to add a new transaction (TX) with $B$,
a new TX block is generated for both $A$ and $B$ and is appended to their respective chains.
The TX block must have a valid hash pointer pointing to the previous block
and a reference\footnote{This is different from the original TrustChain definition found in~\cite{trustchain}.
In there, a TX block has two outgoing edges which are hash pointers to the two parties involved in the transaction.
This work uses one outgoing edge and a reference.} to its \emph{pair} on $B$'s chain.
An example of is shown in \Cref{fig:trustchain-bad}.

@@

If every node follows the rules of TrustChain and we only consider hash pointers and not references,
then every chain effectively forms a singly linked list.
However, if a node violates the rules, then a \emph{fork} may happen.
That is, there may be more than one TX block with a hash pointer pointing back to the same block.
In \Cref{fig:trustchain-bad}, node $b$ (in the middle chain) created two TX blocks that both point to $t_{b, 5}$.
If this is a ledger system it can be seen as a double spend, where the currency accumulated up until $t_{b, 5}$ are spent twice.

@@

\subsection{Consensus protocol}
\label{sec:overview-cons}

@@

The high communication complexity of Byzantine consensus algorithms prohibits us from running it on a large network.
Thus, for every round, we randomly select some node---called facilitators---to collect CP blocks from every other node and use them as the proposal.
The facilitators are elected using a \emph{luck value}, which is computed using $\textsf{H}(\C_r || pk_u)$,
where $\C_r$ is the consensus result 
(which can be seen as the set union of all the CP blocks collected by the facilitators)
in round $r$ and $pk_u$ is the public key of node $u$.
Intuitively, the election is guaranteed to be random 
because the output of a cryptographically secure hash function is unpredictable and $\C_r$ cannot be determined in advance.
Finally, the elected facilitators run a Byzantine consensus algorithm to agree on a set of CP blocks.
The next round begins when the new facilitators are computed from these CP blocks.

@@

The consensus and transaction protocol by themselves do not provide a mechanism to detect forks or other forms of tamperaing.
Thus we need a validation protocol to counteract malicious behaviour.
When a node wish to validate one of its TX, it asks the counterparty for the \emph{agreed fragment} of the TX.
An agreed fragment of a TX is a section of the chain beginning and ending with CP blocks that contains the TX and are in consensus.
Upon the counterparty's response, the node checks that the CP blocks are indeed in consensus,
the hash pointers are valid and his TX is actually in the fragment.
The TX is valid if these conditions are satisfied.
Intuitively, this works because it is hard (because a cryptographically secure hash function is second-preimage resistant)
to create a different chain that begins and ends with the same two CP blocks.

@@

This section and the ones following it give a technical treatment of what the content in System overview.
For notational clarity, we use the following convention (adapted from~\cite{miller2016honey}) for most of this work.
\begin{itemize}
\item Lower case (e.g. $x$) denotes a scalar object or a tuple.
\item Upper case (e.g. $X$) denotes a set or a constant.
\item Sans serif (e.g. $\textsf{fn}(\cdot)$) denotes a function.
\item Monospace (e.g. $\texttt{ack}$) denotes message type.
\end{itemize}
Further, we use $a || b$ to denote concatination of the binary representations of $a$ and $b$.

We assume purely asynchronous channels with eventual delivery.
Thus in no stage of our protocol do we make timing assumptions.
The adversary has fully control of the delivery schedule and the message ordering of all messages.
But they are not allowed to drop messages except for their own\footnote{
    Reliability can be achieved in unreliable networks by resending messages or using some error correction code.
}.

We assume there exist a Public Key Infrastructure (PKI), and nodes are identified by their unique and permanent public key.
Finally, we use the random oracle (RO) model, i.e. calls to the random oracle are denoted by $\textsf{H}: \{0, 1\}^* \rightarrow \{0, 1\}^\lambda$,
where $\{0, 1\}^*$ denotes the space of finite binary strings and $\lambda$ is the security parameter.
Under the RO model, the probability of successfully computing the inverse of the hash function is negligible with respect to $\lambda$~\cite{bellare1993random}.

% TODO big N does not follow our convention
In our model we consider $N$ nodes, which is the population size.
$n$ of them are facilitators, $t$ out of $n$ are malicious and the inequality
$n \ge 3t + 1$ must hold.
This is from the work of Pease, Shostak and Lamport, where they show a network of $n$ nodes cannot reach Byzantine agreement with $t \ge n/3$.
Further, the inequality $N \ge n + t$ must also hold.
This is due to our system design, which becomes clear in~\Cref{sec:consensus-phase}.

Our threat model is as follows. 
We use a restricted version of the adaptive corruption model.
The first restriction is that corrupted node can only change across rounds.
That is, if a round has started, the corrupted nodes cannot changed until the next round.
The second restriction is that the adversary, presumably controlling all the corrupted nodes, is forgetful.
Namely the adversary may learn the internal state such as the private key of a corrupted node,
but if the node recovers, then the adversary must forget the private key.
This is realistic because otherwise the adversary can eventually learn all the private keys and sabotage the system.

\section{Extended TrustChain}
The primary data structure used in our system is the Extended TrustChain.
Each node $u$ has a public and private key pair---$pk_u$ and $sk_u$, and a chain $B_u$.
The chain consist of blocks $B_u = \{ b_{u, i} : i \in \{ 0, \dots, h - 1 \} \}$,
where $b_{u, i}$ is the $i$th block of $u$,
and $h$ is the height of the block (i.e. $h = |B_u|$).
We often use $b_{u, h}$ to denote the latest block.
There are two types of blocks, TX blocks and CP blocks.
If $T_u$ is the set of all TX blocks in $B_u$ and $C_u$ is the set of all CP blocks is $B_u$,
then it must be the case that $T_u \cup C_u = B_u$ and $T_u \cap C_u = \varnothing$.
The notation $b_{u, i}$ is generic over the block type.
We assume there exist a function $\textsf{typeof}: B_u \rightarrow \{ \tau, \gamma \}$ that returns the type of the block,
where $\tau$ represents the TX type and $\gamma$ represents the CP type.

\begin{definition}
\textbf{\emph{Transaction block}}

The TX block is a six-tuple, i.e $t_{u, i} = \langle \textsf{H}(b_{u, i - 1}), seq_u, txid, pk_v, m, sig_u \rangle$.
We describe each item in turn.
\begin{enumerate}
\item $\textsf{H}(b_{u, i - 1})$ is the hash pointer to the previous block.
\item $seq_u$ is the sequence number which should equal $i$.
\item $txid$ is the transaction identifier, it should be generated using a cryptographically secure pseudo-random number generator by the initiator of the transaction.
\item $pk_v$ is the public key of the counterparty $v$.
\item $m$ is the transaction message.
\item $sig_u$ is the signature created using $sk_u$ on the concatination of the binary representation of the five items above.
\end{enumerate}
The fact that we have no constraint on the content of $m$ is in alignment with our design goal---application neutrality.

TX blocks come in pairs.
In particular, for every block 
$$t_{u, i} = \langle \textsf{H}(b_{u, i - 1}), seq_u, txid, pk_v, m, sig_u \rangle$$
there exist one and only one \emph{pair} 
$$t_{v, j} = \langle \textsf{H}(b_{v, j - 1}), seq_v, txid, pk_u, m, sig_v \rangle.$$
Note that the $txid$ and $m$ are the same, and the public keys refer to each other.
Thus, given a TX block, these properties allow us to identify its pair.
\end{definition}

@@

\begin{definition}
\textbf{\emph{Checkpoint block}}

The CP block is a five-tuple, 
i.e. $c_{u, i} = \langle \textsf{H}(b_{u, i-1}), seq_u, \textsf{H}(\C_r), r, sig_u \rangle$,
where $\C_r$ is the consensus result (which we describe in \Cref{sec:consensus-result}) in round $r$, the other items are the same as the TX block definition.
% Note that unlike in our prior work~\cite{implicitconsensus}, CP blocks and TX blocks do not have independent sequence numbers.

The genesis block in the chain must be a CP block in the form of
$c_{u, 0} = \langle \textsf{H}(\bot), 0,  \textsf{H}(\bot), 0, sig_u \rangle$
where $\textsf{H}(\bot)$ can be interpreted as applying the hash function on an empty string.
The genesis block is unique because every node has a unique public and private key pair.
\end{definition}

\begin{definition}
\label{sec:consensus-result}
\textbf{\emph{Consensus result}}

Our consensus protocol runs in rounds as discussed in \Cref{sec:system-overview}.
Every round is identified by a round number $r$, which is incremented on every new round.
The consensus result is a tuple, i.e. $\C_r = \langle r, C \rangle$,
where $C$ is a set of CP blocks agreed by the facilitators of round $r$.
\end{definition}

Here we define an important properties which results from the interleaving nature of CP and TX blocks.
It is used in our validation protocol.
\begin{definition}
\textbf{\emph{Enclosure and agreed enclosure}}

If there exist a tuple $\langle c_{u,a}, c_{u, b} \rangle$ for a TX block $t_{u, i}$,
where 
$$a = \argmin_{k, k < i, \texttt{typeof}(b_{u,k}) = \gamma}(i - k)$$
$$b = \argmin_{k, k > i, \texttt{typeof}(b_{u,k}) = \gamma}(k - i),$$
then $\langle c_{u,a}, c_{u, b} \rangle$ is the enclosure of $t_{u, i}$.
Intuitively, that is the two closest CP blocks that surround $t_{u, i}$.
Note that $c_{u, a}$ is the more recent CP block.
Also, some TX blocks may not have any enclosure, then its enclosure is $\bot$.
Agreed enclosure is the same as enclosure with an extra constraint where the CP blocks must be in some consensus result $\C_r$.

\end{definition}

Note that the ``closest CP blocks'' property must also apply to agreed enclosure. 
Suppose a chain is in the form
$\{c_{i}, c_{i+1}, t_{i+2}, c_{i+3}\}$\footnote{Usually the notation is of the form $c_{u, i}$, but the node identity is not important here so we simplify it to $c_{i}$}
and $c_{i}, c_{i+1}, c_{i+3}$ are in $\C_r, \C_{r+1}, \C_{r+3}$ respectively,
then the agreed enclosure of $t_{i+2}$ is $\langle c_{i+1}, c_{i+3}\rangle$ and cannot be $\langle c_{i}, c_{i+3}\rangle$.

\begin{definition}
\textbf{\emph{Fragment and agreed fragment}}

If the enclosure of some TX block $t_{u, i}$ is $\langle c_{u,a}, c_{u, b} \rangle$,
then its \emph{fragment} $F_{u, i}$ is defined as $\{ b_{u, i} : a \le i \le b \}$.
Similarly, \emph{agreed fragment} has the same definition as fragment but using agreed enclosure.
For convenience, the function $\textsf{agreed\_fragment}(\cdot)$ represents the agreed fragment of some TX block if it exists, otherwise $\bot$.
\end{definition}

@@

Our scalable consensus protocol runs on top of Extended TrustChain.
It uses an unmodified asynchronous subset consensus (ACS) algorithm as the key building block.
The objectives of the protocol are to
    allow honest nodes always make progress (in the form of creating new CP blocks),
    compute correct consensus result in every round
    and have unbiased election of facilitators.
Concretely, we define the necessary properties as follows.
\begin{definition}
\label{def:consensus}
\textbf{\emph{Properties of the consensus protocol}}

$\forall r \in \mathbb{N}$, the following properties must hold.
\begin{enumerate}
    \item \emph{Agreement}:
        If one correct node outputs a set of facilitators $\F_r$,
        then every node outputs $\F_r$
    \item \emph{Validity}:
        If any correct node outputs $\F_r$, then 
            \begin{enumerate}
                \item $|\C_r| \ge N - t$ and each CP in $\C_r$ belong to a different node\footnote{
                $\C_r$ is a tuple but we abuse the notation here by writing $|\C_r|$ to mean the number of CP blocks in the second element of $\C_r$.}.
                \item $\F_r$ must contain at least $n - t$ honest nodes and
                \item $|\F_r| = n$.
            \end{enumerate}
    \item \emph{Fairness}:
        Every node with a CP block in $\C_r$ should have an equal probability of becoming a member of $\F_r$.
    \item \emph{Termination}:
        Every correct node eventually outputs some $\F_r$.
\end{enumerate}
\end{definition}
These properties may look like Byzantine consensus properties (which we describe next in \Cref{sec:acs-background}) but there are subtle differences.
Firstly, they are properties for every node in the network and not just the facilitators.
Secondly, they must be satisfied for all rounds because the whole system falls apart if one of the property cannot be satisfied in one of the rounds.

Before describing the protocol in detail,
we take a brief detour to give background on the ACS algorithm as it is the primary building block.
Then we move on to describe the two phases of our consensus protocol---bootstrap phase and consensus phase.

@@

The best way to explain ACS is to contrast it with the typical Byzantine consensus.
We adapt the description from~\cite[Chapter 17]{podc}.
%The Byzantine consensus problem is a classical problem is distributed systems literature.
%It is first described by Leslie Lamport as the Byzantine general's problem in 1982~\cite{lamport1982byzantine}.
%Byzantine consensus is described as follows (adapted from ).
\begin{definition}
\textbf{\emph{Byzantine consensus}}

@@

ACS is an especially useful primitive for blockchain systems.
It allows any party to propose a value and the result is the set union of all the proposed values by the majority.
Concretely, ACS needs to satisfy the following properties (adapted from~\cite{miller2016honey}).
\begin{definition}
\label{def:acs}
\textbf{\emph{Asynchronous subset consensus}}

@@

ACS has the nice property of censorship resiliance when compared to other consensus algorithms.
For instance, Hyperledger and Tendermint uses a variant of Practical Byzantine Fault Tolerance (PBFT)~\cite{castro1999practical} as their consensus algorithm.
In PBFT, a leader is elected, if the leader is malicious but follows the protocol, then it can selectivly filter transactions.
In contract, every party in ACS are involved in the proposal phase,
and it guarantees that if $n - 2t$ parties propose the same transaction, then it must be in the agreed output.
Thus, if some value is submitted to at least $n - 2t$ nodes, it is guaranteed to be in the consensus result.
For a detailed description of ACS we refer to the HoneyBadgerBFT work~\cite{miller2016honey}.
To understand the remainder of this work, the knowledge of~\Cref{def:acs} is sufficient.

@@

\subsection{Bootstrap phase}
\label{sec:bootstrap}
Now we have all the necessary information to describe our consensus protocol.
As with all distributed systems, there must be a bootstrap phase which sets up the system.

The reader may notice from the system overview (\Cref{sec:system-overview}) that the facilitators are computed from the consensus result,
but the consensus result is agreed by the facilitators.
Thus we have a dependency cycle.
The goal of the bootstrap phase is to give us a starting point in the cycle.

To bootstrap, imagine that there is some bootstrap oracle, that initiates the code on every node.
The code satisfies all the properties in \Cref{def:consensus}.
Namely every node has the same set of valid facilitators $\F_1$ that are randomly choosen.
This concludes the bootstrap phase.
For any future rounds, the consensus phase is used.

@@

\subsection{Consensus phase}
\label{sec:consensus-phase}
The consensus phase begins when $\F_r$ is available to all the nodes.
Note that $\F_r$ indicates the facilitators that were elected using results of round $r$ and are responsible for driving the ACS protocol in round $r + 1$.
The goal is to reach agreement on a set of new facilitators $\F_{r + 1}$ that satisfies the four properties in \Cref{def:consensus}.

There are two scenarios in the consensus phase.
First, if node $u$ is not the facilitator, it sends $\langle \texttt{cp\_msg}, c_{u, h} \rangle$ to all the facilitators.
Second, if the node is a facilitator, it waits until $N - t$ messages of type \texttt{cp\_msg} are received.
% Second if the node is a facilitator, it waits for some duration $D$ and collect messages of type \texttt{cp\_msg}, where $D \gg \Delta$.
Invalid messages are removed, namely blocks with invalid signatures and blocks signed by the same key.
With the sufficient number of \texttt{cp\_msg} messages,
it begins the ACS algorithm and some $\C'_{r + 1}$ should be agreed upon by the end of it.
Duplicates and blocks with invalid signatures are again removed from $\C'_{r+1}$ and we call the final result $\C_{r+1}$.
We have to remove invalid blocks a second time (after ACS) because the adversary may send different CP blocks to different facilitators,
which results in invalid blocks in the ACS output, but not in any of the inputs.

At the core of the consensus phase is the ACS protocol.
While any ACS protocol that satisfies the standard definition will work,
we use a simplification of HoneyBadgerBFT~\cite{miller2016honey} as our ACS protocol
because it is a consensus algorithms designed for blockchain systems.
We do not use the full HoneyBadgerBFT due to the following.
First, the transactions in HoneyBadgerBFT are first queued in a buffer and the main consensus algorithm starts only when the buffer reaches an optimal size.
We do not have an infinite stream of CP blocks, thus buffering is unsuitable.
Second, HoneyBadgerBFT uses threshold encryption to hide the content of the transactions.
But we do not reach consensus on transactions, only CP blocks, so hiding CP blocks is meaningless for us as it contains no transactional information.

@@

Upon receiving $\C_{r + 1}$ and at least $n - f$ valid signatures by some node $u$, $u$ performs two asks.
First, it creates a new CP block using $\textsf{new\_cp}(\C_{r + 1}, r + 1)$ using \Cref{alg:new-cp}.
Second, it computes the new facilitators using $\textsf{get\_facilitator}(\C_{r + 1}, n)$ using \Cref{alg:facilitator},
and updates its facilitator set to the result.
This concludes the consensus phase and brings us back to the beginning of the consensus phase.
